TrueSkill Through Time: the \proglang{Julia}, \proglang{Python} and \proglang{R} packages

### 2021-01-23 article.tex
TTT: Comienzo sección ejemplos

@@

\en{The class \texttt{Rating()} is represented by: the mean (\texttt{mu}) and the deviation (\texttt{sigma}) of the prior Gaussian distribution; the performance standard deviation (\texttt{beta}); and the temporal volatility of the skills (\texttt{gamma}).}
\es{La clase \texttt{Rating()} se representa con cuatro atributos: la media (\texttt{mu}) y el desv\'io (\texttt{sigma}) de la distribuci\'on gaussiana a priori; el dev\'io de los rendimientos (\texttt{beta}); y la volatilidad temporal de las habilidades (\texttt{gamma}).}
%
\begin{lstlisting}[backgroundcolor=\color{white},label=lst:rating, caption=\relax, belowskip=-1.0 \baselineskip, aboveskip=-0 \baselineskip]
\end{lstlisting}
\begin{lstlisting}[backgroundcolor=\color{all}]
mu = 0.0; sigma = 6.0; beta = 1.0; gamma = 0.0
r1 = Rating(mu, sigma, beta, gamma); r2 = Rating(mu, sigma, beta, gamma)
r3 = Rating(mu, sigma, beta, gamma); r4 = Rating(mu, sigma, beta, gamma)
\end{lstlisting}  
%
\en{For efficiency, the prior is stored pre-computed.}
\es{Por cuesti\'on de eficiencia, el prior Gaussiano se guarda precomputado.}
%
\begin{lstlisting}[backgroundcolor=\color{white},label=lst:prior, caption=\relax, belowskip=-1.0 \baselineskip, aboveskip=-0 \baselineskip]
\end{lstlisting}
\begin{paracol}{3}
\begin{lstlisting}[backgroundcolor=\color{julia}]
prior = r1.prior
beta == r1.beta
gamma == r1.gamma
\end{lstlisting}
\switchcolumn
\begin{lstlisting}[backgroundcolor=\color{python}]
prior = r1.prior
beta == r1.beta
gamma == r1.gamma
\end{lstlisting}
\switchcolumn
\begin{lstlisting}[backgroundcolor=\color{r}]
prior = r1$prior
beta == r1$beta
gamma == r1$gamma
\end{lstlisting}
\end{paracol}
%

\en{The class \texttt{Game()} is used to compute the posteriors and the evidence of events such as described in Figure~\ref{fig:generative_model}.}
\es{La clase \texttt{Game()} se utiliza para computar los posteriors y la evidencia de eventos como los descritos en la figura~\ref{fig:generative_model}.}
%
\en{As an example, we initialize a game with two teams of two players.}
\es{A modo de ejemplo, inicializamos una partida con dos equipos de dos jugadores.}
%
\begin{lstlisting}[backgroundcolor=\color{white},label=lst:game, caption=\relax, belowskip=-1.0 \baselineskip, aboveskip=-0 \baselineskip]
\end{lstlisting}
\begin{paracol}{3}
\begin{lstlisting}[backgroundcolor=\color{julia}, belowskip=-0.77 \baselineskip]
team_a = [ r1, r2 ]
team_b = [ r3, r4 ]
teams = [team_a, team_b]
result = [1,0]
\end{lstlisting}
  \switchcolumn
\begin{lstlisting}[backgroundcolor=\color{python}, belowskip=-0.77 \baselineskip]
team_a = [ r1, r2 ]
team_b = [ r3, r4 ]
teams = [team_a, team_b]
result = [1,0]
\end{lstlisting}
   \switchcolumn
\begin{lstlisting}[backgroundcolor=\color{r}, belowskip=-0.77 \baselineskip]
team_a = c(r1, r2)
team_b = c(r3, r4)
teams = list(team_a, team_b)
result = c(1,0)
\end{lstlisting}  
\end{paracol}
\begin{lstlisting}[backgroundcolor=\color{all}]
g = Game(teams, result)
\end{lstlisting}
%
\en{Its attributes, besides the \texttt{teams} and the \texttt{result}, are the approximate (Gaussian) \texttt{likelihoods} and the \texttt{evidence}.}
\es{Sus atributos, adem\'as de los \texttt{teams} y el \texttt{result}, son los \texttt{likelihoods} aproximados (Gaussianos) y la \texttt{evidencia}.}
%
\begin{lstlisting}[backgroundcolor=\color{white},label=lst:game_attributes, caption=\relax, belowskip=-1.0 \baselineskip, aboveskip=-0 \baselineskip]
\end{lstlisting}
\begin{paracol}{3}
\begin{lstlisting}[backgroundcolor=\color{julia}]
teams == g.teams
result == g.result
lhs = g.likelihoods
ev = g.evidence
\end{lstlisting}
  \switchcolumn
\begin{lstlisting}[backgroundcolor=\color{python}]
teams == g.teams
result == g.result
lhs = g.likelihoods
ev = g.evidence
\end{lstlisting}
   \switchcolumn
\begin{lstlisting}[backgroundcolor=\color{r}]
teams == g$teams
result == g$result
lhs = g$likelihoods
ev = g$evidence
\end{lstlisting}  
\end{paracol}
%
\en{In the presence of teams, the observed result is modeled as a direct consequence of the difference of the team performance, $d = \sum_{i \in A} p_i - \sum_{i \in B} p_i$.}
\es{En presencia de equipos, el resultado observado se modela como consecuencias directa de la diferencia de los rendimientos de cada equipo, $d = \sum_{i \in a} p_i) - \sum_{i \in b} p_i)$}
%
\en{In the figure~\ref{fig:modelo_trueskill_2vs2}, we show its graphical factorization.}
\es{En la figura~\ref{fig:modelo_trueskill_2vs2} mostramos su factorizaci\'on gr\'afica.} 
%
\begin{figure}[h!]
  \centering
  \scalebox{.9}{
  \tikz{

@@

%
\begin{comment}

\begin{CodeChunk}
\begin{CodeInput}
R> data("quine", package = "MASS")
\end{CodeInput}
\end{CodeChunk}

This tutorial will guide you through some of typical \textbf{TTT} package application.
Familiarity with Python is assumed, so if you are new to Python there is plenty of online documentation like the official Python documentation
page at \href{http://www.python.org/doc/}{\textcolor{blue}{http://www.python.org/doc/}}.
In addition to the module \textbf{TTT}, the libraries collections, dateutil and mathematics(Gaussian) are also required.
In this section its show some applications, starting with an example of TrueSkill of 1vs1, then an example with a data set, and the last one a first example of Trueskill Through Time.

\subsection{Example 1: TrueSkill 1vs1}
The first examples is the most basic one we figure out. Its one of the first examples in the official python package of \texttt{Trueskill}.
In this case, we consider a game that can be won or lost but not tied, in which two players face each other and we want to update our prior of them knowing the result of it.
For that we start creating an environment of \texttt{TrueSkill}, specifying the parameter \texttt{draw\_probability} to zero, not to consider the possibility of a tie.
Then we define the prior of each player. 
Because in this case we don't know anithing of each, they start with a default prior.
This one can be set up at will.
The parameters of this, is the mean and standard deviation of the skill (mu and sigma), the standard deviation of the performance (beta), and the noise add it in the posterior, as the method suggest (noise).
The result of the game, is in format of a list, with a number for each player, meaning 0 for the one that lost the game, and 1 for the winner.
After this, define the game with the method \texttt{Game}, and calculate this the posterior of the player with the method \texttt{posterior}.
We update the prior of each player with the posterior.

\subsection{Example 4: Estimating the Evolution of the Skill}

%\todo{inline]{poner el ejemplo de seguimiento de la habilidad de un jugador a medida que van pasando los partidos con una funci\'on de evoluci\'on conocida. ejemplo sinte\'etico}

\subsection{Example 5: Performance Analysis}

%% -- Summary/conclusions/discussion -------------------------------------------

>>>>

%
\begin{comment}
\begin{CodeChunk}
\begin{CodeInput}
R> data("quine", package = "MASS")
\end{CodeInput}
\end{CodeChunk}

\en{In this section we show how to use \proglang{Julia}, \proglang{Python} and \proglang{R} packages to solve a single event, a short history of 3 events and the real history of professional ATP tennis players.}
\es{En esta sección mostramos cómo usar los usando los paquetes de\proglang{Julia}, \proglang{Python} y \proglang{R} para resolver un evento, una pequeña historia de 3 eventos y finalmente la historia real de jugadores profesional de tenis ATP.}
%
\en{We will show both TrueSkill and TrueSkill Through Time solutions, the method to obtain individual learning curves, and the a priori prediction of the observed data (i.e. evidence) that serve as a measure to compare models.}
\es{Mostraremos la soluciones TrueSkill y TrueSkill Through Time, el método para obtener curvas de aprendizaje individuales, y la predicción a priori (i.e. evidencia) del dato observado que sirven como medida para comparar modelos.}
%
\en{Finally we will make a time complexity comparison between the three packages presented in this paper and the \texttt{trueskill} package of presented in this paper \proglang{Python}.}
\es{Por último haremos un comparación de complejidad temporal entre los tres paquetes que presentamos en este trabajo y el paquete \texttt{trueskill} de \proglang{Python}.}

\en{The class \texttt{Rating()} is represented by: the mean (\texttt{mu}) and the standard deviation (\texttt{sigma}) of the prior Gaussian distribution; the standard deviation of the performance (\texttt{beta}); and the dinamic factor of the skill (\texttt{gamma}).}
\es{La clase \texttt{Rating()} se representa con cuatro atributos: la media (\texttt{mu}) y el desv\'io estandar (\texttt{sigma}) de la distribuci\'on gaussiana a priori; el dev\'io estanndar de los rendimientos (\texttt{beta}); y el factor din\'amico de la habilidad (\texttt{gamma}).}
%
\begin{lstlisting}[backgroundcolor=\color{white},label=lst:rating, caption=\relax, belowskip=-1.0 \baselineskip, aboveskip=-0 \baselineskip]
\end{lstlisting}
\begin{lstlisting}[backgroundcolor=\color{all}]
mu = 0.0; sigma = 6.0; beta = 1.0; gamma = 0.0
r1 = Rating(mu, sigma, beta, gamma); r2 = Rating(mu, sigma, beta, gamma)
r3 = Rating(mu, sigma, beta, gamma); r4 = Rating(mu, sigma, beta, gamma)
\end{lstlisting}  
%
\en{Ratings have 3 attributes.}
\es{Los rating tienen 3 atributos.}
%
\begin{lstlisting}[backgroundcolor=\color{white},label=lst:prior, caption=\relax, belowskip=-1.0 \baselineskip, aboveskip=-0 \baselineskip]
\end{lstlisting}
\begin{paracol}{3}
\begin{lstlisting}[backgroundcolor=\color{julia}]
prior = r1.prior
beta == r1.beta
gamma == r1.gamma
\end{lstlisting}
\switchcolumn
\begin{lstlisting}[backgroundcolor=\color{python}]
prior = r1.prior
beta == r1.beta
gamma == r1.gamma
\end{lstlisting}
\switchcolumn
\begin{lstlisting}[backgroundcolor=\color{r}]
prior = r1$prior
beta == r1$beta
gamma == r1$gamma
\end{lstlisting}
\end{paracol}
%
\en{For efficiency reasons, the prior is stored as \texttt{Gaussian()} class, which has the mean and standard deviation as attributes.}
\es{Por cuesti\'on de eficiencia, el prior se guarda como clase \texttt{Gaussian()}, la que tiene la media y el desv\'io estandar como atributos.}
%
\begin{lstlisting}[backgroundcolor=\color{white},label=lst:prior, caption=\relax, belowskip=-1.0 \baselineskip, aboveskip=-0 \baselineskip]
\end{lstlisting}
\begin{paracol}{3}
\begin{lstlisting}[backgroundcolor=\color{julia}]
mu1 = r1.prior.mu
sigma1 = r1.prior.sigma
\end{lstlisting}
\switchcolumn
\begin{lstlisting}[backgroundcolor=\color{python}]
mu1 = r1.prior.mu
sigma1 = r1.prior.sigma
\end{lstlisting}
\switchcolumn
\begin{lstlisting}[backgroundcolor=\color{r}]
mu1 = r1$prior$mu
sigma1 = r1$prior$sigma
\end{lstlisting}
\end{paracol}
%
\en{The \texttt{Gaussian()} class does most of the computation of the packages (details at section \ref{sec:Gasussian}).}
\es{La clase \texttt{Gaussian()} realiza la mayor parte del c\'omputo en todos los paquetes (detalles en la secci\'on \ref{sec:Gaussian}).}

\en{The class \texttt{Game()} is used to compute the posteriors and the evidence of events such as described in Figure~\ref{fig:generative_model}.}
\es{La clase \texttt{Game()} se utiliza para computar los posteriors y la evidencia de eventos como los descritos en la figura~\ref{fig:generative_model}.}
\en{As an example, we initialize a game with two teams of two players.}
\es{A modo de ejemplo, inicializamos una partida con dos equipos de dos jugadores.}
\begin{lstlisting}[backgroundcolor=\color{white},label=lst:game, caption=\relax, belowskip=-1.0 \baselineskip, aboveskip=-0 \baselineskip]
\begin{lstlisting}[backgroundcolor=\color{julia}, belowskip=-0.77 \baselineskip]
team_a = [ r1, r2 ]
team_b = [ r3, r4 ]
teams = [team_a, team_b]
result = [1,0]
\begin{lstlisting}[backgroundcolor=\color{python}, belowskip=-0.77 \baselineskip]
team_a = [ r1, r2 ]
team_b = [ r3, r4 ]
teams = [team_a, team_b]
result = [1,0]
\begin{lstlisting}[backgroundcolor=\color{r}, belowskip=-0.77 \baselineskip]
team_a = c(r1, r2)
team_b = c(r3, r4)
teams = list(team_a, team_b)
result = c(1,0)
g = Game(teams, result, p_draw = 0.0)
\en{In the presence of teams, the observed result is modeled as a direct consequence of the difference of the team performance, $d = \sum_{i \in A} p_i - \sum_{i \in B} p_i$ (see details at section \ref{sec:2vs2}).}
\es{En presencia de equipos, el resultado observado se modela como consecuencias directa de la diferencia de los rendimientos de cada equipo, $d = \sum_{i \in a} p_i) - \sum_{i \in b} p_i)$ (detalles en la sección \ref{sec:2vs2}).}
\en{The attributes of the \texttt{Game()} class are the \texttt{teams}, the \texttt{result}, the approximate \texttt{likelihoods} (one for each agent) and the \texttt{evidence}.}
\es{Los atributos de la clase \texttt{Game()} son los \texttt{teams} y el \texttt{result}, los \texttt{likelihoods} aproximados (uno por cada agente) y la \texttt{evidencia}.}
\begin{lstlisting}[backgroundcolor=\color{white},label=lst:game_attributes, caption=\relax, belowskip=-1.0 \baselineskip, aboveskip=-0 \baselineskip]
\begin{lstlisting}[backgroundcolor=\color{julia}, belowskip=-0.77 \baselineskip]
teams == g.teams
result == g.result
lhs = g.likelihoods
ev = g.evidence
ev = round(ev, digits=3)
\begin{lstlisting}[backgroundcolor=\color{python}, belowskip=-0.77 \baselineskip]
teams == g.teams
result == g.result
lhs = g.likelihoods
ev = g.evidence
ev = round(ev, 3)
\begin{lstlisting}[backgroundcolor=\color{r}, belowskip=-0.77 \baselineskip]
teams == g$teams
result == g$result
lhs = g$likelihoods
ev = g$evidence
ev = round(ev, 3)
print(ev)
> 0.5
\en{The evidence is $0.5$, indicating that the a priori probability of the observed result was the same for both teams.}
\es{La evidencia es $0.5$, indicando que la probabiilidad a priori del resultado observado era la misma para ambos equipos.}
\en{The posteriors of each agent are Gasussian that can be computed from the likelihoods and priors by calling a method.}
\es{Los posteriors de cada agente son gaussianas que pueden computarse de los likelihoods y priors llamando a un método.}
\en{The output has the same order in which the teams and agents were loaded.}
\es{La salida tiene el mismo orden en el que fueron cargados los equipos y los agentes.}
\begin{lstlisting}[backgroundcolor=\color{white},label=lst:game_attributes, caption=\relax, belowskip=-1.0 \baselineskip, aboveskip=-0 \baselineskip]
\begin{lstlisting}[backgroundcolor=\color{julia}, belowskip=-0.77 \baselineskip]
posts = posteriors(g)
print(posts[1][1], ev)
\begin{lstlisting}[backgroundcolor=\color{python}, belowskip=-0.77 \baselineskip]
posts = g.posteriors
print(posts[0][0], ev)
\begin{lstlisting}[backgroundcolor=\color{r}, belowskip=-0.77 \baselineskip]
posts = g$posteriors()
print(posts[[1]][[1]], ev)
> Gaussian(mu=2.361, sigma=5.516)

\subsection{Environment}

\en{Before solving a history of events it is necessary to review the set of hyperparameters of the model (\texttt{environment}): the mean \texttt{mu} and standard deviation \texttt{sigma} of the priors; the standard deviation of the performances \texttt{beta}; the dynamic factor \texttt{gamma}; the draw probability \texttt{p\_draw}; and the convergence conditions \texttt{epsilon} and \texttt{iter}.}
\es{Antes de resolver una historia de eventos es necesario revisar el conjunto de hiperparámetros del modelo (\texttt{environment}): la media \texttt{mu} y el desvío estandar \texttt{sigma} de los priors; el desvío estandar de los desempeños \texttt{beta}; el factor dinámico \texttt{gamma}; la probabilidad de empate \texttt{p\_draw}; y las condiciones de convergencia \texttt{epsilon} e \texttt{iter}.}
%
\begin{lstlisting}[backgroundcolor=\color{all}]
env = Environment(mu=0., sigma=6., beta=1., gamma=.05, p_draw=0., epsilon=1e-6, iter=10)
env == Environment()
\end{lstlisting}

\subsection{Single event}
The first examples is the most basic one we figure out.
Its one of the first examples in the official python package of \texttt{Trueskill}.
In this case, we consider a game that can be won or lost but not tied, in which two players face each other and we want to update our prior of them knowing the result of it.
For that we start creating an environment of \texttt{TrueSkill}, specifying the parameter \texttt{draw\_probability} to zero, not to consider the possibility of a tie.
Then we define the prior of each player. 
Because in this case we don't know anithing of each, they start with a default prior.
This one can be set up at will.
The parameters of this, is the mean and standard deviation of the skill (mu and sigma), the standard deviation of the performance (beta), and the noise add it in the posterior, as the method suggest (noise).
The result of the game, is in format of a list, with a number for each player, meaning 0 for the one that lost the game, and 1 for the winner.
After this, define the game with the method \texttt{Game}, and calculate this the posterior of the player with the method \texttt{posterior}.
We update the prior of each player with the posterior.

\subsection{Example 4: Performance Analysis}

@@

\subsection{\en{Graphical factorization}\es{Factorizaci\'on gr\'afica}}\label{sec:2vs2}

\en{In the figure~\ref{fig:modelo_trueskill_2vs2}, we show its graphical factorization.}
\es{En la figura~\ref{fig:modelo_trueskill_2vs2} mostramos su factorizaci\'on gr\'afica.} 
%
\begin{figure}[h!]
  \centering
  \scalebox{.9}{
  \tikz{

@@
